---
title: "Classification: LDA and QDA"
author: "YOUR NAME HERE"
format: 
  html:
    code-fold: true
    code-line-numbers: true
    code-tools: true
    self-contained: true
editor: visual
execute:
  message: false
---

## Setup

Declare your libraries:

```{r}
#| label: libraries-r
#| include: false
library(tidyverse)
library(tidymodels)
library(glmnet)
library(discrim)
```


# LDA

```{r}
ins <- read_csv("https://www.dropbox.com/s/bocjjyo1ehr5auz/insurance.csv?dl=1")

ins <- ins %>%
  mutate(
    smoker = factor(smoker)
  ) %>%
  drop_na()
```



```{r}
lda_mod <- discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")
```


```{r}
lda_fit_1 <- lda_mod %>%
  fit(smoker ~ charges, data = ins)

lda_fit_1$fit %>% summary()
```


```{r}
preds <- lda_fit_1 %>% predict(ins)

ins <- ins %>%
  mutate(
    pred_smoker = preds$.pred_class
  )

ins %>%
  accuracy(truth = smoker,
           estimate = pred_smoker)
```


```{r}
lda_fit_2 <- lda_mod %>%
  fit(smoker ~ charges + age, data = ins)

lda_fit_2
```




```{r}
lda_fit_2$fit$scaling
```

```{r}
ins %>%
  ggplot(aes(x = charges, y = age, color = smoker)) +
  geom_point()
```

```{r}
lda_fit_2

my_slope = lda_fit_2$fit$scaling[1]/(-1*lda_fit_2$fit$scaling[2])
```

```{r}

ins %>%
  ggplot(aes(x = charges, y = age, color = smoker)) +
  geom_point() +
  geom_abline(aes(slope = my_slope, intercept = 0))
```

#### Your turn:

Use cross-validation to find the best LDA model to predict smoker status.

How does it compare to the Logistic Regression and KNN approaches?

# Quadratic Discriminant Analysis

#### Code:

```{r}
qda_mod <- discrim_regularized(frac_common_cov = 0) %>% 
             set_engine('klaR') %>% 
             set_mode('classification')
```


#### Your turn:

Use Cross-Validation to find the best QDA model to predict smoker status.

How does it compare to the LDA, Logistic Regression, and KNN approaches?

# Metrics

https://yardstick.tidymodels.org/articles/metric-types.html

